---
title: "My title"
author: 
  - Sinan Ma
thanks: "Code and data are available at: LINK."
date: today
date-format: long
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(tidymodels)
library(modelsummary)
library(nflverse)
```

# Data

```{r}
#| echo: false
#| message: false
#| warning: false

nfl_2023 <- read_csv(here::here("data/analysis_data.csv"))
nfl_2023_train <- nfl_2023 |>
  filter(week<=9)
nfl_2023_test <- nfl_2023 |>
  filter(week>9)

```


```{r}
#| label: tbl-model
#| tbl-cap: "Explanatory Model of NFL Quarterback Passing EPA"
#| echo: false


first_model_tidymodels <-
  linear_reg() |>
  set_engine(engine = "lm") |>
  fit(
    passing_epa ~ passing_yards + sacks + interceptions,
    data = nfl_2023_train
  )

second_model_tidymodels <-
  linear_reg() |>
  set_engine(engine = "lm") |>
  fit(
    passing_epa ~ passing_tds + sacks + attempts,
    data = nfl_2023_train
  )


modelsummary(list(model_1 = first_model_tidymodels,
                  model_2 = second_model_tidymodels))
```

```{r}
#| echo: false

predicted_results <- first_model_tidymodels |>
  predict(new_data = nfl_2023_test) |>
  cbind(nfl_2023_test) |>
  select(week, recent_team, passing_epa, .pred) |>
  arrange(week, recent_team) |>
  rename(predicted_passing_epa = .pred)
```


# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

\begin{align} 
y_i | \mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 x_{1i} + \beta_2 x_{2i} + \beta_3 x_{3i} + \beta_4 x_{4i} + \beta_5 x_{5i}
\end{align}

where $y_i$ refers to the total expected points added on pass attempts and sacks; $\mu_i$ is the mean of the normal distribution for observation $i$, which is the predicted passing EPA; $\sigma$ is the standard deviation of the normal distribution across all quarterbacks' EPA. $x_{1i}$ to $x_{5i}$ correspond to the predictors for observation $i$, which in the context of the model would be `passing_yards`, `sacks`, `interceptions`, `passing_tds`, and `attempts`; $\beta_1$ to $\beta_5$ are the coefficients for the predictor variables.



### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Result
```{r}
#| label: fig-result
#| tbl-cap: "Comparison of Actual vs Predicted Passing EPA"
#| echo: false

predicted_results |>
  ggplot(aes(x = passing_epa, y = predicted_passing_epa)) +
  geom_point(alpha = 0.5) + 
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "blue") +
  labs(
    title = "Actual vs Predicted Passing EPA",
    x = "Actual Passing EPA",
    y = "Predicted Passing EPA"
  ) +
  theme_minimal() + 
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```



# Discussion
xxx


\newpage


# References


